Laplacian envelope

### 2021-08-12 Notes/laplace.tex
diagonal constraints

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
%avoid double superscripts
\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  \Delta f\, \eq \e^{rt\frg} \w ,
\end{aligned}
\end{equation}
%
where \(\Delta f = f\pot - f\dep \).
Combining the expressions in \cref{eq:overlapwt} allows us to write the numerator of \cref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
                 {\sqrt{1-(\Delta f)^2(\eq \w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eq_+=\eq_-\).
We can perform the maximisation in two steps.
First maximise the numerator at fixed \(\eq \w\).
Then maximise the ratio \wrt \(\eq\w\).
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of \(\MMd^\nu\) as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
  \qquad
  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
  \sum_j \frgm_{ij} &=0, &
  \sum_j \encm_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

The usual fundamental matrix identities will hold provided we use \cref{eq:lshift} as the transition matrix and we note that the \( \arow \) we used should be interpreted as \( \arow(s)\) and the ``original'' \( \arow \) is \( \arow + s \eq \equiv \shift{\arow}\) and the ``original'' \( \frac{1}{\tau_0} \) is \( \frac{1}{\tau_0} + s \equiv \frac{1}{\shift{\tau}_0}\).
We could rewrite \cref{eq:lfundrowsum} as
%
\begin{equation}\label{eq:lshiftrowsum}
\begin{aligned}
  \fund(s) \onev &= {\shift{\tau}_0}\onev, \qquad &
  \shift{\MM}(s) \fund(s) &= \onev \eq - \I, \\ 
  \shift{\arow} \fund(s) &= \eq,  &
  \fund(s) \shift{\MM}(s) &= \shift{\tau} \onev \shift{\arow} - \I.
\end{aligned}
\end{equation}
%

This means that the \( \fptb(s) \) have all the same properties as in the case \(s=0\).
This includes non-negativity and all of the triangle inequalities.
Identities involving mean first-passage times and the transition matrix will also hold, provided that \cref{eq:lshift} is used as the transition matrix.
In particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
\end{equation}
%
is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}\).
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
\end{equation}
%
We can arrange the states in order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing \(\etwm_i(s)\) is the same as the order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
The \(\etwm_i(s)\) can still be used when \(\w\) takes more than two values.

@@

We can also express this in terms of the shifted process in \cref{eq:larea}.
We can choose how to distribute the shift across \(\MMd \pot\) and \(\MMd \dep\).
Two choices from a continuum%
\footnote{We could choose \( \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
are:
%
\begin{equation}\label{eq:lshiftpd}
  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
  \qquad \text{or} \qquad
  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{r} (\onev \eq - \I).
\end{equation}
%
The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} (f\pot - f\dep) (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
Either way:
%
\begin{equation}\label{eq:lareashift}
  \hat{A}\prn{s; \MMd^\nu} = \hat{A}\prn{0; \shift{\MMd}^\nu(s)}.
\end{equation}
%
In other words, the laplace transform of the SNR is the area under the SNR for the shifted process.

@@

In analogy with the case \(s=0\) in \cite{Lahiri2013synapse}, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    {c}_k(s) &= \sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
    {a}_j(s) &= \sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
    \thbm_i(s) &= \sum_j \fptbm_{ij}(0)\, \eqm_j {a}_j(s), \\
  \end{aligned}
    \implies
    \hat{A}(s) 
      % = \sum_k \eqm_k {c}_k(s) \wm_k
      = \mathbf{c}^\dag(s) \w
      = \eq \mathbf{a}(s).
\end{equation}
%
If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) wlg.),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).

Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\MMdm^\mu_{mn}}
     = f^\mu \eqm_m \prn{ r \brk{\thbm_m(s) - \thbm_n(s)}
     +\prn{r {c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
\end{equation}
%
The term with \(\thb(s)\) comes from differentiating \(\eq\), the term with \(c(s)\) comes from differentiating \(\etw(s)\) and the term with \((-1)^\mu\) comes from differentiating \(\enc\) in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
We define the matrix \(\mathbf{X}(s) = \fund(0) \enc \fund(s)\).
Given a four-index array \(H_{ijkl}\), we define \(H\trans_{ijkl} = H_{klij}\).
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{alignedat}{3}
    R_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \circ \partial \frgm_{kl}}} 
    &= \
    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
    &&&+ \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
          \brc{\etwm_k(s) - \etwm_l(s)} \\
    &&&+ \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
          \brc{\etwm_k(s) - \etwm_l(s)},  \\
  S_{ijkl} = \smash{\pdiff{^2 \hat{A}}{\frgm_{ij} \partial \encm_{kl}} }
  &= \
    &&\eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\etwm_k(s) - \etwm_l(s)} \\
    &&&+ \eqm_k \brc{\fundm_{ki}(s) - \fundm_{li}(s)} \brc{\etwm_i(s) - \etwm_j(s)},
\end{alignedat}
\end{equation}
%
where \(R\) contains only those terms where \( \frgm_{ij} \) precedes \( \frgm_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
The other terms sum to \(R\trans\).

Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
  \pdiff{^2\hat{A}(s)}{\MMdm^\mu_{ij} \partial \MMdm^\nu_{kl}} =
     f^\mu f^\nu [R + R\trans + (-1)^\mu S\trans + (-1)^\nu S]_{ijkl}.
\end{equation}
%
This is useful in numerical maximisation.

@@

\begin{figure}[tbp]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=7]{envelope.pdf}}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=13]{envelope.pdf}}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=21]{envelope.pdf}}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=27]{envelope.pdf}}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=36]{envelope.pdf}}
  \item\aligntop{\includegraphics[width=0.45\linewidth,page=50]{envelope.pdf}}
  \end{myenuma}
  \caption[Optimal models]{The models that constitute the numeric envelope in \cref{fig:envnum}.
  The plot to the left of each frame indicates the memory curve of the model shown to the right in green, with the numeric envelope for models with the serial topology in orange and the proven envelope, \eqref{eq:env}, in blue.
  Above the left plot we have a graphical display of the model, described with more detail to the right.
  The area of each node is proportional to the steady state probability in that state and the thickness of the red and blue edges are proportional to the transition probabilities under potentiation and depression.
  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
  See the video for the full set of models.}\label{fig:envvid}
\end{figure}

>>>>

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  \Delta f\, \eq \e^{rt\frg} \w ,
\end{aligned}
\end{equation}
%
where \(\Delta f = f\pot - f\dep \).
Combining the expressions in \cref{eq:overlapwt} allows us to write the numerator of \cref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
                 {\sqrt{1-(\Delta f)^2(\eq \w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with \(f\pot=f\dep\) or \(\eq_+=\eq_-\).
We can perform the maximisation in two steps.
First maximise the numerator at fixed \(\eq \w\).
Then maximise the ratio \wrt \(\Delta \pr = \eq\w\).
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of \(\MMd^\nu\) as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
  \qquad
  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of \(\frg\) and \(\enc\) as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq 2f^\mu, \\
  \sum_j \frgm_{ij} &=0, &
  \sum_j \encm_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

The usual fundamental matrix identities will hold provided we use \cref{eq:lshift} as the transition matrix and we note that the \( \arow \) we used should be interpreted as \( \arow(s)\) and the ``original'' \( \arow \) is \( \arow + s \eq \equiv \shift{\arow}\) and the ``original'' \( \frac{1}{\tau_0} \) is \( \frac{1}{\tau_0} + s \equiv \frac{1}{\shift{\tau}_0}\).
We could rewrite \cref{eq:lfundrowsum} as
%
\begin{equation}\label{eq:lshiftrowsum}
\begin{aligned}
  \fund(s) \onev &= \shift{\tau}_0 \onev, \qquad &
  \shift{\MM}(s) \fund(s) &= \onev \eq - \I, \\ 
  \shift{\arow} \fund(s) &= \eq,  &
  \fund(s) \shift{\MM}(s) &= \shift{\tau}_0 \onev \shift{\arow} - \I.
\end{aligned}
\end{equation}
%

This means that the \( \fptb(s) \) have all the same properties as in the case \(s=0\).
This includes non-negativity and all of the triangle inequalities.
Identities involving mean first-passage times and the transition matrix will also hold, provided that \cref{eq:lshift} is used as the transition matrix.
In particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
\end{equation}
%
is independent of the starting state \(i\), but now \(\eta(s) = \tr\fund(s) - \shift{\tau}_0 \).
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
\end{equation}
%
We can arrange the states in order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing \(\etwm_i(s)\) is the same as the order of decreasing \(\eta^{+}_i(s)\) or increasing \(\eta^{-}_i(s)\).
The \(\etwm_i(s)\) can still be used when \(\w\) takes more than two values.

@@

We can also express this in terms of the shifted process in \cref{eq:larea}.
We can choose how to distribute the shift across \(\MMd \pot\) and \(\MMd \dep\).
Two choices from a continuum%
\footnote{We could choose \( \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s g^\nu}{r f^\nu} (\onev \eq - \I) \), for any \(g^\nu \) with \( \sum_\nu g^\nu = 1 \).
It shifts \(\frg\) correctly and also shifts \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta g (\onev \eq - \I) \).
The two possibilities in \cref{eq:lshiftpd} come from \(g^\nu = \frac{1}{2}\) and \(g^\nu = f^\nu\) respectively.}
are:
%
\begin{equation}\label{eq:lshiftpd}
  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{2 r f^\nu} (\onev \eq - \I)
  \qquad \text{or} \qquad
  \shift{\MMd}^\nu(s) = \MMd^\nu + \frac{s}{r} (\onev \eq - \I).
\end{equation}
%
The first choice leaves \( \shift{\enc} = \enc \), so it only affects the \(\etw\) piece of \cref{eq:larea}.
The second choice does not leave \(\enc\) invariant (in fact \( \shift{\enc}(s) = \enc + \frac{s}{r} \Delta f (\onev \eq - \I) \)), but because \( \eq (\onev \eq - \I) = 0 \) the shift does not make any contribution to \cref{eq:larea}.
Either way:
%
\begin{equation}\label{eq:lareashift}
  \hat{A}\prn{s; \MMd^\nu} = \hat{A}\prn{0; \shift{\MMd}^\nu(s)}.
\end{equation}
%
In other words, the laplace transform of the SNR is the area under the SNR for the shifted process.

@@

In analogy with the case \(s=0\) in \cite{Lahiri2013synapse}, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    {c}_k(s) &= r\sum_{ij} \eqm_i \encm_{ij} \prn{\fptbm_{ik}(s)-\fptbm_{jk}(s)}, \\
    {a}_j(s) &= r\sum_{k} \encm_{jk} (\etwm_{j}(s) - \etwm_{k}(s)), \\
    \thbm_i(s) &= \sum_j \fptbm_{ij}(0)\, \eqm_j {a}_j(s), \\
  \end{aligned}
    \implies
    \hat{A}(s) 
      % = \sum_k \eqm_k {c}_k(s) \wm_k
      = \mathbf{c}^\dag(s) \w
      = \eq \mathbf{a}(s).
\end{equation}
%
If we regard the \(\wm_k\) as optimisable parameters (limited to the range \( [-1,1] \) wlg.),s the optimal choice is clearly \( \wm_k = \sgn(c_k) \).

Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\MMdm^\mu_{mn}}
     = f^\mu \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
     +\prn{{c}_m(s) + (-1)^\mu} \brk{\etwm_m(s) - \etwm_n(s)} }.
\end{equation}
%
The term with \(\thb(s)\) comes from differentiating \(\eq\), the term with \(c(s)\) comes from differentiating \(\etw(s)\) and the term with \((-1)^\mu\) comes from differentiating \(\enc\) in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
We define the matrix \(\mathbf{X}(s) = r\fund(0) \enc \fund(s)\).
Given a four-index array \(H_{ijkl}\), we define \(H\trans_{ijkl} = H_{klij}\).
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{alignedat}{2}
    R_{ijkl} &= \smash{\pdiff{\hat{A}}{\frgm_{ij} \circ ; \frgm_{kl}}} &
    &=
    \eqm_i \brc{\fundm_{ik}(0) - \fundm_{jk}(0)} \brc{\thbm_k(s) - \thbm_l(s)} \\
    &&&\hphantom{=} {}
    + \eqm_i \brc{{X}_{ik}(s) - {X}_{jk}(s)} 
          \brc{\etwm_k(s) - \etwm_l(s)} \\
    &&& \hphantom{=} {}
    + \eqm_i {c}_i(s) \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} 
          \brc{\etwm_k(s) - \etwm_l(s)},  \\
  S_{ijkl} &= \smash{\pdiff{\hat{A}}{\encm_{ij} ; \frgm_{kl}} } &
  &= \eqm_k \brc{\fundm_{ki}(0) - \fundm_{li}(0)} \brc{\etwm_i(s) - \etwm_j(s)} \\
    &&& \hphantom{=} {}
    + \eqm_i \brc{\fundm_{ik}(s) - \fundm_{jk}(s)} \brc{\etwm_k(s) - \etwm_l(s)},
\end{alignedat}
\end{equation}
%
where \(R\) contains only those terms where \( \frgm_{ij} \) precedes \( \frgm_{kl} \) in the matrix products from the expression for \( \hat{A} \) in \cref{eq:larea}.
The missing terms sum to \(R\trans\).

Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
  \pdiff{^2\hat{A}(s)}{\MMdm^\mu_{ij} \partial \MMdm^\nu_{kl}} =
     f^\mu f^\nu [R + R\trans + (-1)^\mu S + (-1)^\nu S\trans]_{ijkl}.
\end{equation}
%
This is useful in numerical maximisation.

@@
